---
title: "Aarhus air raid shelters in 1944"
author: "Manuel Ø. Mørch"
date: "2026-05-01"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


```{r Reading the data, message=FALSE, warning=FALSE}
library(tidyverse)
library(sf)

# First we will read the relevant data from the MELICA dataset.The data can be downloaded here: https://github.com/MELICA-AU/shelter-data/blob/master/output_data/BDG_long.rds

tilflugtsrum <- readRDS("BDG_long.rds")

#And now we will take a look at the data: 
glimpse(tilflugtsrum)

#In order to only keep those from 1944, i will filter the dataframe by the year of construction:

tilflugtsrum_1944 <- tilflugtsrum %>% 
  filter(Location_startdate == 1944)
 
```



# First modern map with markers for the air raid shelters:

```{r mapping the air raid shelters on a modern map, warning=FALSE}
library(leaflet)
library(htmlwidgets)



# And now to add the other columns as popups:

tilflugtsrum_map <- leaflet(tilflugtsrum_1944) %>%
  addTiles() %>%
  addMarkers(popup = paste("Capacity:", tilflugtsrum_1944$Capacity, "<br>",
                           "Location certainty:", tilflugtsrum_1944$Certain_location, "<br>",
                           "Accuracy:", tilflugtsrum_1944$Accuracy, "<br>",
                           "Notes:", tilflugtsrum_1944$Notes))

saveWidget(tilflugtsrum_map, "Tilflugtsrum.html", selfcontained = TRUE)

tilflugtsrum_map

```

# Heatmap of the air raid shelters:

```{r air raid shelters on modern heatmap, warning=FALSE}
library(leaflet.extras)

tilflugtsrum_heatmap <- leaflet(tilflugtsrum_1944) %>% 
  addTiles() %>%
  addHeatmap(intensity = 5, # you can specify an intensity variable if you have one
    blur = 20        # adjust the blur radius
  )
tilflugtsrum_heatmap

saveWidget(tilflugtsrum_heatmap, "TilflugtsrumHeatmap.html", selfcontained = TRUE)

```
These maps are quite nice, however they show us the bunkers' locations on a moderne map.
This is an issue because the landscape has changed a lot since 1944, and we want to see what areas were protected in 1944.
However, since the MELICA dataset contains shelters outside of the Touristmap's coverage, we will add both the Touristmap and OpenStreetMap as layers, so that we can toggle the modern map on and off based on whether the area of interest is covered by the Touristmap or not.

# Adding the historical map for appropriate spatial analysis of the historical period.

```{r Historical with markers, warning=FALSE}

touristmap_markers <- leaflet(tilflugtsrum_1944) %>%
   setView(lng = 10.2039, lat = 56.1629, zoom = 13) %>%
  # Adding the OpenStreetMap layer for a modern map
   addTiles() %>%
  # Adding the historical Map layer
  addTiles(urlTemplate = "TouristMap_Tiles/{z}/{x}/{y}.png",
    options = tileOptions(minZoom = 0, maxZoom = 18, tms = TRUE), 
    group = "TouristMap (1942-1946)") %>%
  # Adding the markers with popups for the shelter information from the MELICA dataset
  addMarkers(
    popup = paste(
      "Capacity:", tilflugtsrum_1944$Capacity, "<br>",
      "Location certainty:", tilflugtsrum_1944$Certain_location, "<br>",
      "Accuracy:", tilflugtsrum_1944$Accuracy, "<br>",
      "Notes:", tilflugtsrum_1944$Notes))%>%
  # Adding a controlpanel to toggle the historical map layer on and off
  addLayersControl(
    overlayGroups = c("TouristMap (1942-1946)"),
    options = layersControlOptions(collapsed = FALSE))

touristmap_markers


saveWidget(touristmap_markers, "Touristmap_markers.html", selfcontained = TRUE)
```


```{r Historical map with heatmap, warning=FALSE}
touristmap_heatmap <- leaflet(tilflugtsrum_1944) %>% 
   setView(lng = 10.2039, lat = 56.1629, zoom = 13) %>%
  # Adding the OpenStreetMap layer for a modern map
  addTiles() %>%
  addTiles(
    urlTemplate = "TouristMap_Tiles/{z}/{x}/{y}.png",
    options = tileOptions(minZoom = 0, maxZoom = 18, tms = TRUE), 
    group = "TouristMap (1942-1946)") %>%
  # Adding the heatmap layer for the shelter locations
  addHeatmap(intensity = 5,
    blur = 20) %>%
   # Adding a controlpanel to toggle the historical map layer on and off
  addLayersControl(
    overlayGroups = c("TouristMap (1942-1946)"),
    options = layersControlOptions(collapsed = FALSE))

touristmap_heatmap

saveWidget(touristmap_heatmap, "Touristmap_heatmap.html", selfcontained = TRUE)
```

